---
title: "USMLE Data Analysis"
author: "Lisa Levoir"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format: 
  html:
        theme: yeti
        #code-fold: true
        code-tools: true
        html-math-method: katex
        toc: true
        toc-depth: 3
        fig-width: 13
        fig-height: 10
        toc-title: "Contents"
        number-sections: true
        self-contained: true
        self-contained-math: true
        smooth-scroll: true
        fontsize: 0.8em
        title-block-banner: true
        citation-location: margin
        include-after-body: graph_fold.html
editor: visual
engine: knitr
---

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(plotly)
library(dplyr)
library(tidyr)
library(knitr)
library(table1) #Create HTML Tables of Descriptive Statistics https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html
#library(OMTM1) #https://github.com/schildjs/OMTM1/
library(Hmisc)
library(rms) # Regression Modeling Strategies by Frank https://cran.r-project.org/web/packages/rms/index.html
library(modelsummary) #Summary Tables and Plots for Statistical Models and Data: Beautiful, Customizable, and Publication-Ready https://cran.r-project.org/web/packages/modelsummary/index.html
library(scales) # The scales packages provides the internal scaling infrastructure used by ggplot2, and gives you tools to override the default breaks, labels, transformations and palettes. https://scales.r-lib.org
library(viridis) #colors
library(cowplot) #allows me to use plotgrid
library(gridExtra) #adding tables to plots
library(visdat) #shows missing data
library(GGally) #makes pairs plots

setwd("/Users/lisalevoir/BIOS7351_Collab/data_project2") #this line I would need to run in the console
knitr::opts_knit$set(root.dir = "/Users/lisalevoir/BIOS7351_Collab/github/BIOS_Collaboration/project_2_analysis") #now I set global options for knitting, I also had to toggle global options > R Markdown > evaluate chunks in current directory


#import the data
dat <- read.csv("/Users/lisalevoir/BIOS7351_Collab/data_project2/combined_data_203.csv") # baseline information + days from randomization to an event

#to compare that the merging went as expected
VUMSdat <- read.csv("/Users/lisalevoir/BIOS7351_Collab/data_project2/DATA_VUMS.csv")
HMSdat <- read.csv("/Users/lisalevoir/BIOS7351_Collab/data_project2/DATA_HMS.csv")
UVAdat <- read.csv("/Users/lisalevoir/BIOS7351_Collab/data_project2/DATA_UVA.csv")
```

### Setup

Criteria to exclude students who most likely took a scored exam:

1.  Any PhD students (n = 2)

2.  Any 5th year program students (n = 19)

3.   M4 students at Vanderbilt (n = 5)

4.  Students who did not complete either Step survey (n = 2)

5.  Students who specifically stated they took a scored Step 1 (n=1)

Based on our criteria we would exclude record IDs: 

-   VUSM: 23, 26, 39, 40, 54, 3, 8, 12, 49, 60, 62, 64 

-   HMS: 1, 21, 28, 30, 34, 37, 39, 41, 44, 47, 49, 61 

-   UVA: 33, 47, 80, 81, 83 

```{r data_cleaning, warning=FALSE, message=FALSE}
#the list of IDs we decided as a group to exclude
excludeVU <- c(23, 26, 39, 40, 54, 3, 8, 12, 49, 60, 62, 64)
excludeHMS <- c(1, 21, 28, 30, 34, 37, 39, 41, 44, 47, 49, 61)
excludeUVA <- c(33, 47, 80, 81, 83)

'%!in%' <- function(x,y)!('%in%'(x,y)) #make a way to use the not in command
VU_in <- filter(VUMSdat, record_id  %!in%  excludeVU)
H_in <- filter(HMSdat, record_id  %!in%  excludeHMS)
UVA_in <- filter(UVAdat, record_id  %!in%  excludeUVA)

#now I want to select the columns I'd like to include for all of my analysis (so they're in the proper order for a cbind). this will be relatively easy to come back to edit later, if needed. 

#first, remember to include a school identifier
VU_in[,"schoolid"] <- "VU"
UVA_in[,"schoolid"] <- "UVa"
H_in[,"schoolid"] <- "HMS"

took_step1_firstVU <-
  VU_in %>% select(
    starts_with("uworld_percent_step1_1"),
    starts_with("amboss_percent_step1_1"),
    starts_with("length_step1_1"),
    starts_with("practicetest_step1_1"),
    starts_with("full_test_practice_step1_1"),
    starts_with("push_step1_1"),
    starts_with("push_practice_test_step1_1"),
    starts_with("push_nbme_practice_score_step1_1"),
    starts_with("push_uw_practice_score_step1_1"),
    starts_with("final_nbme_practice_score_step1_1"),
    starts_with("final_uw_practice_score_step1_1"),
    starts_with("score_step1_1"),
    "schoolid"
  ) #note, I removed ""number_other_courses_step1_1" and starts_with("other_courses_step1_1___1) because these questions were not on the other school surveys

took_step1_firstUVA <-
  UVA_in %>% select(
    starts_with("uworld_percent_step1_1"),
    starts_with("amboss_percent_step1_1"),
    starts_with("length_step1_1"),
    starts_with("practicetest_step1_1"),
    starts_with("full_test_practice_step1_1"),
    starts_with("push_step1_1"),
    starts_with("push_practice_test_step1_1"),
    starts_with("push_nbme_practice_score_step1_1"),
    starts_with("push_uw_practice_score_step1_1"),
    starts_with("final_nbme_practice_score_step1_1"),
    starts_with("final_uw_practice_score_step1_1"),
    starts_with("score_step1_1"),
    "schoolid"
  )

took_step1_firstH <-
  H_in %>% select(
    starts_with("uworld_percent_step1_1"),
    starts_with("amboss_percent_step1_1"),
    starts_with("length_step1_1"),
    starts_with("practicetest_step1_1"),
    starts_with("full_test_practice_step1_1"),
    starts_with("push_step1_1"),
    starts_with("push_practice_test_step1_1"),
    starts_with("push_nbme_practice_score_step1_1"),
    starts_with("push_uw_practice_score_step1_1"),
    starts_with("final_nbme_practice_score_step1_1"),
    starts_with("final_uw_practice_score_step1_1"),
    starts_with("score_step1_1"),
    "schoolid"
  )


took_step1_first<- rbind(took_step1_firstH, took_step1_firstUVA, took_step1_firstVU) #great! this is clean data. I can follow a similar process for the other data. This cleaned data set is less relevant since we can't run logistic regression since all participants passed, but it is a good proof of the workflow.

visdat::vis_miss(took_step1_first) #shows a lot of missing data
```

```{r data_cleaning_forstep2, warning=FALSE, message=FALSE}
#these are the variables I can actually analyze
table(VU_in$score_step2_2) 
table(VU_in$score_step2_1)

took_step2_VU <-
  VU_in %>% select(
    starts_with("uworld_percent_step2"),
    starts_with("amboss_percent_step2"),
    starts_with("length_step2"),
    starts_with("practicetest_step2"),
    starts_with("full_test_practice_step2"),
    starts_with("practice_score_step2"),
    starts_with("score_step2"),
    starts_with("target_score_step2"),
    "schoolid"
  ) #note, I removed ""number_other_courses_step1_1" and starts_with("other_courses_step1_1___1) because these questions were not on the other school surveys

took_step2_UVA <-
  UVA_in %>% select(
    starts_with("uworld_percent_step2"),
    starts_with("amboss_percent_step2"),
    starts_with("length_step2"),
    starts_with("practicetest_step2"),
    starts_with("full_test_practice_step2"),
    starts_with("practice_score_step2"),
    starts_with("score_step2"),
    starts_with("target_score_step2"),
    "schoolid"
  )

took_step2_H <-
  H_in %>% select(
    starts_with("uworld_percent_step2"),
    starts_with("amboss_percent_step2"),
    starts_with("length_step2"),
    starts_with("practicetest_step2"),
    starts_with("full_test_practice_step2"),
    starts_with("practice_score_step2"),
    starts_with("score_step2"),
    starts_with("target_score_step2"),
    "schoolid"
  )

took_step2general <- rbind(took_step2_H, took_step2_UVA, took_step2_VU)

took_step2_first <- took_step2general %>% select(ends_with("_1"), "schoolid")
took_step2_second <- took_step2general %>% select(ends_with("_2"), "schoolid")
visdat::vis_miss(took_step2_first) #looks like a school specific trend with who is answering what question...
visdat::vis_miss(took_step2_second) #looks like a school specific trend with who is answering what question...

# mod1 <- lm(score_step2_2 ~ ., data = took_step2_second) #so this is nonsense right now but just seeing what would happen
# summary(mod1)

tf = function(x,y) {
  ct <- cor.test(x, y)
  sprintf("Corr: %0.2f, p: %0.3f\n95%% CI: (%0.2f - %0.2f) \n n pairs = %.0f", 
          ct$estimate, ct$p.value, ct$conf.int[1], ct$conf.int[2], sum(!is.na(x), !is.na(y))/2)
}  #https://stackoverflow.com/questions/75068871/add-p-value-and-or-95-ci-in-the-text-of-a-correlation-matrix-pairs-plot




class(took_step2_second$practice_score_step2_2) <- "integer"
class(took_step2_second$score_step2_2) <- "integer"
ggpairs(took_step2_second, title="Pairs plots", upper=list(continuous=wrap("statistic",text_fn=tf,title=NULL, sep=NULL)), progress = FALSE) #well this means it's time to take a break


```


# another note
this was another tactic I was trying but frustratingly, the class of different columns was changed between surveys. So far, I have abandoned this technique in favor of the above, where I am affirmativly selecting the columns (but I still may run into the different class issue)
```{r}
#try dropping problematic columns and then do bind_rows
# removeextracolsH <-
#   H_in %>% select(
#     -c(
#       "vusmhms_usmle_trends_survey_timestamp",
#       "vusmhms_usmle_trends_survey_complete"
#     ),
#     contains("number_other_courses")
#   )
# setdiff(names(H_in), names(UVA_in))
# setdiff(names(removeextracolsH), names(UVA_in))
# 
# removeextracolsUVA <-
#   UVA_in %>% select(
#     -contains("number_other_courses"),
#     -c(
#       "vusmuva_usmle_trends_survey_timestamp",
#       "vusmuva_usmle_trends_survey_complete"
#     )
#   )
# setdiff(names(removeextracolsUVA), names(removeextracolsH)) #confirming the cols match now
# 
# removeextracolsVU <- VU_in %>% select(-contains("number_other_courses"), -contains("other_courses"))
# setdiff(names(removeextracolsUVA), names(removeextracolsVU)) #confirming the cols match now
# 
# class(removeextracolsH$practicetest_step1_1) <- "integer" #this column was coded as a character so it wouldn't merge previously
# bind_rows(list(removeextracolsVU, removeextracolsH), .id = "id")

```

### remember to profile missingness for students who conducted only one Step Survey
