---
title: "BayesSurvival Vignette Example for Presentation"
author: "Alexis Fleming"
output:
  html_document:
    toc: true
    df_print: paged
  pdf_document: default
---

## Setup
```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir=file.path("/Users/alexisfleming/Survival/Final Presentation/Vignette Example"))

# load libraries
library(tidyverse)
library(bayesplot)
library(BayesSurvival)
library(survival)
library(survminer)
library(simsurv)
library(ggfortify)
library(GGally)

# set seed for reproducibility
set.seed(7323)
```


## Example from README
```{r README example}
hazard.true <- function(t,x, betas, ...){1.2*(5*(t+0.05)^3 - 10*(t+0.05)^2 + 5*(t+0.05) ) + 0.7}
sim.df <- data.frame(id = 1:1000)
df <- simsurv(x = sim.df, maxt = 1, hazard = hazard.true)

bs <- BayesSurv(df, "eventtime", "status")

```


## Example from Vignette with dependent Gamma prior
Data load and Clean
```{r clean data}
# the cancer df is from the survival package
# Survival in patients with advanced lung cancer from the North Central Cancer Treatment Group. Performance scores rate how well the patient can perform usual daily activities.

cancer$status[cancer$status == 1] <- 0 #censored
cancer$status[cancer$status == 2] <- 1 #died
table(cancer$status)
```

Interesting note: The use of 1/2 for alive/dead instead of the usual 0/1 is a historical footnote. For data contained on punch cards, IBM 360 Fortran treated blank as a zero, which led to a policy within the section of Biostatistics to never use "0" as a data value since one could not distinguish it from a missing value. The policy became a habit, as is often the case; and the 1/2 coding endured long beyond the demise of punch cards and Fortran.


Plot Survival Curves with the Cancer dataset
```{r dep Gamma prior surv curves}
# Example with dependent Gamma prior
res_dep <- BayesSurv(df = cancer, #our data frame
                     time = "time", #name of column with survival/censoring times
                     event = "status", #name of column with status indicator
                     prior = "Dependent") #use dependent Gamma prior

# Plot the survival curve from the BayesSurv function
km_plot_bayes_dep <- 
  PlotBayesSurv(bayes.surv.object = res_dep,
                object = "survival",
                color = "olivedrab4",
                legend=F,
                ylab = "Survival", 
                xlab="Time",
                plot.title = "Bayesian Survival (Dep. Prior) & KM Curve")

# the previous way that's been demonstrated in class to create a survival curve
km_surv <- survfit(Surv(time=cancer$time , event=cancer$status) ~ 1,
                   data=cancer,
                   type="kaplan-meier")

#dataframe of time and survival 
df.km <- data.frame(t = km_surv$time, km = km_surv$surv)
# Plot survival K-M curve
km = ggsurv(km_surv, main = 'KM Survival Curve', 
            back.white = TRUE, size.est = 1, size.ci = 1, 
            surv.col = 'royalblue4')
# autoplot(km_surv) +
#  labs(x = "\n Survival Time (Days) ", y = "Survival Probabilities \n",
#  title = "Survival Times Of \n Advanced Lung Cancer Patients \n") +
#  theme(plot.title = element_text(hjust = 0.5),
#  axis.title.x = element_text(face="bold", colour="#FF7A33", size = 12),
#  axis.title.y = element_text(face="bold", colour="#FF7A33", size = 12),
#  legend.title = element_text(face="bold", size = 10))

#Add KM curve (+ CI) to Bayesian survival curve
km_plot_bayes_dep <- 
  km_plot_bayes_dep + 
  geom_step(data = km$data, aes(x = time, y = surv), col = 'royalblue4', linewidth = 1) +
  geom_line(data = km$data, aes(x = time, y = up), col = 'royalblue4', linewidth = 1, linetype = 81) +
  geom_line(data = km$data, aes(x = time, y = low),col = 'royalblue4', linewidth = 1, linetype = 81)
km_plot_bayes_dep 
```

Plot the Cumulative Hazard with the Cancer dataset
```{r dep Gamma prior cumhaz curves}
# Bayesian output of Cumulative Hazard curve
cumhaz_plot_bayes_dep <- 
  PlotBayesSurv(bayes.surv.object = res_dep,
                object = "cumhaz",
                color = "olivedrab4",
                ylab = "Cumulative hazard",
                xlab = "Time",
                legend = F,
                plot.title = 'Bayesian Cumulative Hazard Curve (Dep. Prior)')

# Cumulative hazard non-Bayes
ggsurvplot(km_surv,
           fun="cumhaz", 
           xlab="Days", 
           ylab="Cumulative Hazard")

# from lab 3:
cumhaz_df <- list(time=km_surv$time, cumhaz = -log(km_surv$surv)) %>% as_tibble()
cumhaz_curve <- 
  ggplot(data = cumhaz_df, aes(x=time,y=cumhaz)) + 
  geom_step(col = 'royalblue4', linewidth = 1)

# Overlay of both plots
overlay_cumhaz_plots <- 
  cumhaz_plot_bayes_dep + 
  geom_step(data = cumhaz_df, 
            aes(x=time,y=cumhaz),
            col = 'royalblue4', 
            linewidth = 1)
overlay_cumhaz_plots


# Bayes Plot inst. hazard
PlotBayesSurv(bayes.surv.object = res_dep,
              object = "hazard",
              plot.title = "Hazard")

#Fit a Cox PH model
model_fit_cph <- coxph(Surv(time=cancer$time , event=cancer$status) ~ 1,
                     data=cancer)
#Use basehaz() to produce estimates of the cumulative hazard
# when instant_haz() is used below, basehaz() is called to get these cumulative hazards
cumhaz<-survival::basehaz(model_fit_cph, center=FALSE)

#load the supplemental function
source("inst_haz.R")

#use instant_haz(coxph.fit) to produce estimates of the baseline instantaneous hazard
instanthaz <- instant_haz(model_fit_cph)

# plot non-Bayes
ggplot(data=instanthaz, aes(x=time,y=inst_haz))+
  geom_smooth()

# overlay both plots - this doesn't seem right?
PlotBayesSurv(bayes.surv.object = res_dep,
              object = "hazard",
              plot.title = "Hazard") + 
  geom_smooth(data=instanthaz, aes(x=time,y=inst_haz), se=F)


```


## Example from Vignette with independent Gamma prior
Plot Survival Curves with the Cancer dataset
```{r indep Gamma prior surv curves}
# Example with independent Gamma prior
res_indep <- BayesSurv(df = cancer, #our data frame
          time = "time", #name of column with survival/censoring times
          event = "status", #name of column with status indicator
          prior = "Independent") #use independent Gamma prior

# Plot the survival curve from the BayesSurv function using indep. Gamma prior
km_plot_bayes_indep <- 
  PlotBayesSurv(bayes.surv.object = res_indep,
                object = "survival",
                color = "olivedrab4",
                legend=F,
                ylab = "Survival", 
                xlab="Time",
                plot.title = "Bayesian Survival (Indep. Prior) & KM Curve")

#Add KM curve (+ CI) to Bayesian survival curve
km_plot_bayes_indep <- 
  km_plot_bayes_indep + 
  geom_step(data = km$data, aes(x = time, y = surv), col = 'royalblue4', linewidth = 1) +
  geom_line(data = km$data, aes(x = time, y = up), col = 'royalblue4', linewidth = 1, linetype = 81) +
  geom_line(data = km$data, aes(x = time, y = low),col = 'royalblue4', linewidth = 1, linetype = 81)
km_plot_bayes_indep 
```

Plot the Cumulative Hazard with the Cancer dataset
```{r indep Gamma prior cumhaz curves}

# Bayesian output of Cumulative Hazard curve
cumhaz_plot_bayes_indep <- 
  PlotBayesSurv(bayes.surv.object = res_indep,
                object = "cumhaz",
                color = "olivedrab4",
                ylab = "Cumulative hazard",
                xlab = "Time",
                legend = F,
                plot.title = 'Bayesian Cumulative Hazard Curve (Indep. Prior)')

# Cumulative hazard non-Bayes
cumhaz_curve <-
  ggsurvplot(km_surv,
             fun="cumhaz", 
             xlab="Days", 
             ylab="Cumulative Hazard")

# from lab 3:
cumhaz_df <- list(time=km_surv$time, cumhaz = -log(km_surv$surv)) %>% as_tibble()
cumhaz_curve <- 
  ggplot(data = cumhaz_df, aes(x=time,y=cumhaz)) + 
  geom_step(col = 'royalblue4', linewidth = 1)

# Overlay of both plots
overlay_cumhaz_plots <- 
  cumhaz_plot_bayes_indep + 
  geom_step(data = cumhaz_df, 
            aes(x=time,y=cumhaz),
            col = 'royalblue4', 
            linewidth = 1)
overlay_cumhaz_plots

PlotBayesSurv(bayes.surv.object = res_indep,
              object = "hazard",
              plot.title = "Hazard")

```



